---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


The code used to unzip the original file is included (so it will be fully "reproducible").

1. Unzip file
2. Read CSV file (with all default parameters)
3. Convert the date to appropriate R format using **as.Date** function

```{r Init, echo=TRUE}
library(lubridate)
library(ggplot2)
library(dplyr)

unzip("activity.zip")
ActivityDS=read.csv("activity.csv")
ActivityDS$date=as.Date(ActivityDS$date)
```

## What is mean total number of steps taken per day?

We can use the function **tapply** to calculate the total number of steps (The result is printed per day)

```{r Step1, echo=TRUE}
Totals<-tapply(ActivityDS$steps,as.factor(ActivityDS$date),sum,na.rm=TRUE)
Totals
```

Then we can plot a histogram of the total number of steps taken each day, and calculate mean and median of the total number of steps taken per day:

```{r Step2, echo=TRUE}
hist(Totals,main="Histogram Total Steps")
mean(Totals,na.rm=TRUE)
median(Totals,na.rm=TRUE)
```

## What is the average daily activity pattern?

We can make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r Step3, echo=TRUE}
Avg_5m<-tapply(ActivityDS$steps,as.factor(ActivityDS$interval),mean,na.rm=TRUE)
plot(x=names(Avg_5m) , y=Avg_5m,type='l',main="Steps by each interval")
```

And then we can find Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps, and, if we divide this number by 60 we can have the hour:


```{r Step4, echo=TRUE}
Avg_5m[Avg_5m==max(Avg_5m)]
as.numeric(names(Avg_5m[Avg_5m==max(Avg_5m)]))/60
```

## Imputing missing values

1. Notice there are many missing values in STEPS 

```{r Step5, echo=TRUE}
sum(is.na(ActivityDS$steps))
```

2. We can choose the following strategy: fill with 0 the missing values (because it make sense that it is 0, maybe the person is inactive,so he/she didn't move)

3. Creating  a new Data Set with 0 in missing values

```{r Step6, echo=TRUE}
ActivityDS2=ActivityDS
ActivityDS2$steps[is.na(ActivityDS2$steps)] = 0
```

4. We will repeat the same calculations with 0 in missing values:

```{r Step7, echo=TRUE}
Totals2<-tapply(ActivityDS2$steps,as.factor(ActivityDS2$date),sum,na.rm=FALSE)
hist(Totals2)

mean(Totals2,na.rm=TRUE)
median(Totals2,na.rm=TRUE)
```

Note: The mean and median of the total steps are not different because in the 1st scenario, the NA sums as 0. So with this strategy, the sum of each day is exactly the same.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r Step8, echo=TRUE}
ActivityDS2$dayWeek=wday(ActivityDS2$date,week_start = 1)
ActivityDS2$WeekEndFlag[ActivityDS2$dayWeek>=5]="Weekend"
ActivityDS2$WeekEndFlag[ActivityDS2$dayWeek<5]="Weekday"

Totals3<-aggregate(ActivityDS2$steps, list(ActivityDS2$interval,
                                           ActivityDS2$WeekEndFlag), mean)

names(Totals3)=c("interval","WeekFlag","Steps")

ggplot(data = Totals3, aes(interval, Steps)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color="steelblue") + 
  labs(title = "Average steps by 5-min interval",
       subtitle = "(groupping by Weekday  /  Weekend)",
       y = "Average steps", x = "") + 
  facet_wrap(~ WeekFlag)

```

It looks like the average steps in weekends are larger in weekends than in a regular weekday.







